---
title: "International Mathematical Olympiad (IMO) Data Analysis: Data Merging Process"
subtitle: "Combining Country Results, Individual Performance, and Competition Timeline Data"
author: "Zhongyue Lin"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: show
    theme: cosmo
    output-file: "index.html"
---

```{r setup, message=FALSE, warning=FALSE}
#| label: setup

# Load required libraries
library(tidyverse)
library(curl)

# Set proxy settings
Sys.setenv(http_proxy = "http://127.0.0.1:1087")
Sys.setenv(https_proxy = "http://127.0.0.1:1087")

# Function to read IMO data
read_imo_data <- function(file_name, integer_cols, character_cols, date_cols = c()) {
  # Construct URL
  url <- paste0("https://raw.githubusercontent.com/Zhongyuelin/Deliverable01-Collecting-raw-data-/main/IMO_data/", file_name)
  
  # Set up curl handle
  h <- new_handle()
  handle_setopt(h, ssl_verifypeer = FALSE)
  temp <- tempfile()
  
  # Download data
  curl_download(url, temp, handle = h)
  
  # Set column types
  col_types <- cols(.default = col_double())
  col_types$cols[integer_cols] <- map(integer_cols, ~ col_integer())
  col_types$cols[character_cols] <- map(character_cols, ~ col_character())
  col_types$cols[date_cols] <- map(date_cols, ~ col_date())
  
  # Read and return data
  data <- read_csv(temp, col_types = col_types)
  unlink(temp)
  return(data)
}

# Read country results data
cat("Reading country results data...\n")
country_results <- read_imo_data(
  "country_results.csv",
  integer_cols = c("year", "team_size_all", "team_size_male", "team_size_female",
                   "p1", "p2", "p3", "p4", "p5", "p6",
                   "awards_gold", "awards_silver", "awards_bronze", "awards_honorable_mentions"),
  character_cols = c("country", "leader", "deputy_leader"))

# Read individual results data
cat("Reading individual results data...\n")
individual_results <- read_imo_data(
  "individual_results.csv",
  integer_cols = c("year", "p1", "p2", "p3", "p4", "p5", "p6", "total", "individual_rank"),
  character_cols = c("contestant", "country", "award"))

# Read timeline data
cat("Reading timeline data...\n")
timeline <- read_imo_data(
  "timeline.csv",
  integer_cols = c("edition", "year", "countries", "all_contestant", 
                   "male_contestant", "female_contestant"),
  character_cols = c("country", "city"),
  date_cols = c("start_date", "end_date"))

cat("All datasets successfully loaded!\n")
```

## Data Merging Process

### 1. Prepare Individual Datasets

First, we prepare each dataset for merging by creating summary statistics and selecting relevant variables.

```{r data-prep}
#| label: data-prep

# Prepare country performance data
country_performance <- country_results %>%
  group_by(year, country) %>%
  summarise(
    team_score = sum(p1 + p2 + p3 + p4 + p5 + p6, na.rm = TRUE),
    total_medals = sum(awards_gold + awards_silver + awards_bronze, na.rm = TRUE),
    team_size = team_size_all,
    .groups = 'drop'
  )

# Prepare individual performance data
individual_performance <- individual_results %>%
  group_by(year, country) %>%
  summarise(
    avg_individual_score = mean(total, na.rm = TRUE),
    top_rank = min(individual_rank, na.rm = TRUE),
    num_contestants = n(),
    .groups = 'drop'
  )

# Prepare timeline data
timeline_summary <- timeline %>%
  select(year, host_country = country, host_city = city, 
         total_countries = countries, total_contestants = all_contestant)
```

### 2. Merge Datasets

Now we merge all prepared datasets into a comprehensive dataset.

```{r merge-data}
#| label: merge-data

# Merge all datasets
comprehensive_imo_data <- country_performance %>%
  left_join(individual_performance, by = c("year", "country")) %>%
  left_join(timeline_summary, by = "year") %>%
  mutate(
    is_host = country == host_country
  ) %>%
  arrange(year, country)

# Create directory and save merged data
dir.create("MergedData", showWarnings = FALSE)
write_csv(comprehensive_imo_data, "MergedData/comprehensive_imo_data.csv")

cat("Merged data has been saved successfully!\n")
```

## Dataset Description

The merged dataset combines information from three different sources and contains the following information:

### 1. Team Performance Data
- Year and country identification
- Team total scores (sum of all problem scores)
- Number of medals won (total of gold, silver, and bronze)
- Team size information

### 2. Individual Performance Data
- Average individual scores per country
- Best ranking achieved by country
- Number of contestants per country

### 3. Competition Context
- Host country and city information
- Total number of participating countries
- Total number of contestants

### Dataset Structure

```{r dataset-structure}
#| label: dataset-structure

# Display structure of merged dataset
cat("Merged Dataset Structure:\n")
str(comprehensive_imo_data)

# Display first few rows
cat("\nSample of Merged Data:\n")
head(comprehensive_imo_data) %>%
  knitr::kable()
```

## Conclusion

This data merging process has successfully:

1. Combined three distinct IMO datasets into a comprehensive dataset
2. Preserved key performance metrics at both team and individual levels
3. Maintained competition context information
4. Created a structured dataset suitable for further analysis

The merged dataset provides a foundation for analyzing:
- Historical trends in IMO performance
- Relationships between team and individual achievements
- Host country performance patterns
- Participation trends over time

All merged data has been saved in the MergedData directory and is ready for further analysis.